---
title: "Rapid Impact Analysis of B 1.1.7 Variant on the Spread of SARS-CoV-2 in North Carolina"
author: Michael DeWitt
bibliography: references.bib
date: 31 January, 2021
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
header-includes:
   - \usepackage{float}
output:
    pdf_document
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r load-data}

```

```{r effects}

```

For correspondence: michael.dewitt@conehealth.com

## Abstract

**Background:** Individual level data 

**Method:** We explored 

**Results:** We found 

**Conclusions:** Population level surveillance data supports 

# Method

## Data

This analysis considered effective reproduction number data generated for the state of North Carolina using the @epinow2 R package following the methods specified by @rtwebsite.
Data on S-gene target failure (SGTF) observed in the United Kingdom and made available in the analysis @davies2020 were used in order to estimate proportion of anticipated B1.1.7 variant circulating.


## Statistical analysis


# Results



```{r scatter, echo=FALSE, fig.width=10, fig.height=10, fig.cap = ""}

```



```{r effects-tab, results = "asis"} 

```

# Discussion


# References

<div id = 'refs'></div>


